Towards Pandemic-Scale Ancestral Recombination Graphs of SARS-CoV-2

### 2023-06-01 paper.tex
Update author section

\title{Towards Pandemic-Scale Ancestral Recombination Graphs of SARS-CoV-2}
\author[1]{Shing H. Zhan}
\author[2,5]{Katherine Eaton}
\author[3,5]{Anastasia Ignatieva}
\author[1,5]{Benjamin Jeffery}
\author[1,5]{Duncan S. Palmer}
\author[1,5]{Yan Wong}
\author[2]{Carmen L. Murall}
\author[4]{Sarah P. Otto}
\author[1,6]{Jerome Kelleher}
\affil[1]{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford, Oxford, United Kingdom}
\affil[2]{National Microbiology Laboratory, Public Health Agency of Canada, Canada}
\affil[3]{Department of Statistics, University of Oxford, Oxford, United Kingdom}
\affil[4]{Department of Zoology and Biodiversity Research Centre, University of British Columbia, Vancouver, British Columbia, Canada}
\affil[5]{Preliminary alphabetic ordering}
\affil[6]{Correspondence. E-mail: jerome.kelleher@bdi.ox.ac.uk}

@@

% Paragraph 4, what do we say about the divergence times?
Figure~\ref{fig:recomb_mrcas} shows that there is a
broad range in divergence times among recombinant parents.
We can see that the majority of recombinants arise
from parents that diverged 10 to 30 weeks
previously. There is also a peak of recombinants that diverged
${\sim}95$ weeks prior to the event, which correspond to recombinants whose
parental common ancestor traces to early 2020. We would expect, for example
Delta-Omicron recombinants to fall into this group, hence it can be
useful to classify the breakpoints further by the VoC combinations of their
parent lineages. Considering only the Alpha, Delta, and Omicron VoC classes,
such a classification reveals that the majority of breakpoints have two Delta or
two Omicron parents, and that Omicron and Delta are the variants associated
with the most recombination (Figure~\ref{fig:recomb_mrcas_voc_breakdown}).
This may reflect either sampling intensity, the
prevalence of cases (which increases the
chance of coinfection and recombination), or possible heterogeneity in
recombination probabilities among lineages.

>>>>

\title{Towards Pandemic-Scale Ancestral Recombination Graphs of SARS-CoV-2}
\author[1]{Shing H. Zhan}
\author[3$\star$]{Anastasia Ignatieva}
\author[1$\star$]{Yan Wong}
\author[2]{Katherine Eaton}
\author[1]{Benjamin Jeffery}
\author[1]{Duncan S. Palmer}
\author[2]{Carmen L. Murall}
\author[4]{Sarah P. Otto}
\author[1$\dagger$]{Jerome Kelleher}
\affil[1]{Big Data Institute, Li Ka Shing Centre for Health
Information and Discovery, University of Oxford, Oxford, United Kingdom}
\affil[2]{National Microbiology Laboratory, Public Health Agency of Canada, Canada}
\affil[3]{Department of Statistics, University of Oxford, Oxford, United Kingdom}
\affil[4]{Department of Zoology and Biodiversity Research Centre,
University of British Columbia, Vancouver, British Columbia, Canada}
\affil[$\star$]{Joint second author}
\affil[$\dagger$]{Correspondence: jerome.kelleher@bdi.ox.ac.uk}

@@

% Paragraph 4, what do we say about the divergence times?
Figure~\ref{fig:recomb_mrcas} shows that there is a
large range in divergence times among recombinant parents, with a broad
spread of times from about 10 to 80 weeks prior to the recombination event
and a minor additional peak of recombinants that diverged
${\sim}95$ weeks, corresponding to recombinants whose
parental common ancestor traces to early 2020. We would expect, for example
Delta-Omicron recombinants to fall into this group, hence it can be
useful to classify the breakpoints further by the VoC combinations of their
parent lineages. Considering only the Alpha, Delta, and Omicron VoC classes,
such a classification reveals that the majority of breakpoints have two Delta or
two Omicron parents, and that Omicron and Delta are the variants associated
with the most recombination (Figure~\ref{fig:recomb_mrcas_voc_breakdown}).
This may reflect either sampling intensity, the
prevalence of cases (which increases the
chance of coinfection and recombination), or possible heterogeneity in
recombination probabilities among lineages.
